---
title: "brain weights"
author: "mcdevitt"
date: "09 avr 2017"
output:
  html_document:
    highlight: tango
    theme: united
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

```

### Homework 12 - Problem Set 2 - Brain weights

```{r brains}

home_dir <- "~/_smu/_src/brain_weights/"
setwd(home_dir)
data_dir <- "./data"
setwd(data_dir)

brains <- read.csv("case0902.csv", stringsAsFactors = FALSE)

setwd(home_dir)

names(brains) <- tolower(names(brains))
brains$species <- as.factor(brains$species)

brains$log_brain <- log(brains$brain)
brains$log_body <- log(brains$body)
brains$log_litter <- log(brains$litter)
brains$log_gestation <- log(brains$gestation)

```

***  

#### ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
### __Step 1 - Plot the data__  

```{r}
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	basic plot of data sets
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

brains$body_group <- cut(brains$log_body, c(-Inf, -2, 0, 2, 4, 6, Inf))
levels(brains$body_group) <- c("<-2", "-2t0", "0-2", "2-4", "4-6", "6+")

p <- ggplot(brains, aes(body, brain))
p + geom_point(aes(color = factor(body_group)))

p <- ggplot(brains, aes(log_body, log_brain))
p + geom_point(aes(color = factor(body_group)))

p <- ggplot(brains, aes(log_litter, log_brain))
p + geom_point(aes(color = factor(body_group)))

p <- ggplot(brains, aes(log_gestation, log_brain))
p + geom_point(aes(color = factor(body_group)))

```


***  

#### ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
### __Steps 2 & 3:  Build a model & Fit the model__  

```{r}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	linear model
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

fit_1 = lm(log_brain ~ log_litter + log_gestation + body_group, data = brains)

summary(fit_1)

pred_1 = data.frame(x = brains$log_litter, y = predict(fit_1, brains), body_group = brains$body_group)

ggplot(brains) +
	geom_point(aes(log_litter, log_brain, color = body_group)) +
	geom_line(data = pred_1, aes(x, y, color = body_group))

pred_2 = data.frame(x = brains$log_gestation, y = predict(fit_1, brains), body_group = brains$body_group)
ggplot(brains) +
	geom_point(aes(log_gestation, log_brain, color = body_group)) +
	geom_line(data = pred_2, aes(x, y, color = body_group))
```

***  

#### ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
#### __Step 4: Provide a residual plot, studentized residual plot, histogram of residuals and qq plot of residuals__  

```{r}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	residuals vs. fitted values
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

library(broom)
df <- augment(fit_1)
ggplot(df, aes(x = .fitted, y = .resid)) + geom_point()


# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	Studentized residuals
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

library(MASS)
sresid <- studres(fit_1)

hist(sresid, freq = FALSE,
   main="Distribution of Studentized Residuals")




```

***  

#### ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
#### __Step 5: if the fit assessed in Step 4 is sufficient, interpret each coefficient in the model.__  

The model accounts for six different grouping of body size.  
For body size grouping as follows :  
A - log(size) < -2 --> size < 0.14 grams  
B - log(size) -2-0 --> size < 1.0 grams  
C - log(size) 0-2 --> size  < 7.39 grams  
D - log(size) 2-4 --> size  < 54.6 grams  
E - log(size) 4-6 --> size  < 403 grams  
F - log(size) 6+ -->  size  > 403 grams  

Then interpretation of model for each body group size is as follows :  

A	brain size = 	litter^0.89 * gestation^1.95  
B	brain size = 	0.33	* litter^0.89 * gestation^1.95  
C	brain size = 	1.60	* litter^0.89 * gestation^1.95  
D	brain size = 	3.94	* litter^0.89 * gestation^1.95  
E	brain size = 	9.68	* litter^0.89 * gestation^1.95  
F	brain size = 	17.12	* litter^0.89 * gestation^1.95  
  

#### ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  

#### BONUS 1pt:  How many degrees of freedom were used to estimate the error term (MSE)? 
#### BONUS 1pt:  What is the estimate of the MSE?  

***  

#### __Residual standard error: 0.51 on 87 degrees of freedom__  

***  
***  

#### ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
end-of-document  
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
